Advice on describing Bayesian analysis of neutron and X-ray reflectometry

### 2022-11-11 src/tex/ms.tex
variance -> variability
First draft following review

% Define document class
\documentclass[reprint,superscriptaddress,aps,amsmath,linenumbers]{revtex4-2}

@@

Reflectometry analysis can be described, in the most simplistic terms, as a comparison and refinement of a model based on some parameters, $\mathbf{x}$, to reproduce some reflectivity data set, $\mathbf{D}$. 
This refinement process involves comparing the model to the data and calculating some goodness-of-fit value or likelihood, $p(\mathbf{D} | \mathbf{x})$, and modifying the model to optimize the goodness-of-fit or maximise the likelihood.
A commonly used goodness-of-fit parameter is the $\chi^2$ parameter which is found as \cite{nelson_refnx_2019}, 
%
\begin{equation}
    \chi^2 = \sum_{q=q_{\text{min}}}^{q_{\text{max}}}{\bigg[\frac{(R(q) - R(q)_{\text{m}})}{\sigma_R(q)}\bigg]^2}, 
\end{equation}
%
where, $R(q)$ and $R(q)_{\text{m}}$ are the measured and modelled reflectivity at a given $q$, the measured momentum transfer vector, and $\sigma_R(q)$ is the uncertainty associated with the measured reflectivity at each $q$ point. 
Under an assumption of normally distributed residuals $R(q) - R(q)_{\text{m}} \sim \mathcal{N}(0, \sigma_R(q))$, the likelihood is related to the $\chi^2$ variable in the following way:
%
\begin{equation}
    \ln[p(\mathbf{D} | \mathbf{x})] = -\frac{1}{2} \bigg(\chi^2 + \sum_{q=q_{\text{min}}}^{q_{\text{max}}}\ln{\big[2\pi\sigma_R(q)^2\big]}\bigg).
    \label{equ:likelihood}
\end{equation}
%
The input for this refinement process is the model and some initial parameter values, which may be an absolute value or some parameter range, depending on the refinement algorithm.
While for some Bayesian sampling process, the input is a probability distribution for each parameter, known as the prior.
The output is a set of values for $\mathbf{x}$, potentially with associated error bars, where these are present they typically describe a standard deviation from the mean of a Gaussian probability distribution. 
A Bayesian sampling process gives a probability distribution, the posterior, the defines the relative likelihood of different values of each parameter, from this we can report statistical measures, e.g. mode/mean/median.
This process implicitly assumes that the data is completely reduced, accounting for all experimental parameters, uncertainties are accurately described and the model can accurately describe the data.

@@

The optimised parameters from the minimisation algorithm, which depend on the particular algorithm, often include some statistical uncertainty.
This uncertainty comes from an assumption of normally distributed parameters \cite{bevington_data_2002}, however, Bayesian sampling approaches make no assumption of an underlying statistical distribution.
How these statistical uncertainties are found is beyond the scope of this work, but it is important to acknowledge that this uncertainty typically assumes that the probability distribution of the parameter is Gaussian in nature. 
This probability distribution is either the partial likelihood or posterior, the latter when some prior is included and Bayes theorem is applied. The posterior describes our understanding of the parameter values given the data that was measured.
When Bayesian modelling is used and the prior is included, the posterior probability is found as, 
%
\begin{equation}
  p(\mathbf{x} | \mathbf{D}) \propto p(\mathbf{D} | \mathbf{x}) p(\mathbf{x}).
  \label{equ:bayes}
\end{equation}
%
Therefore, when Bayesian modelling is performed, the priors and likelihood are of fundamental importance to the results that are obtained (the posterior), and any scientific conclusions that are drawn.

@@

Bayes theorem (given in Equation~\ref{equ:bayes}) consists of the product of the prior and the likelihood. 
The former describes our current understanding of the parameters before we conduct some experiments, while the latter describes how well the data is described by the model parameters. 
Therefore, we must state how the likelihood for a given model and data is calculated (giving the analysis package used, version number, and if not the default the likelihood option), in particular, because although Equation~\ref{equ:likelihood} is a common approach, it assumes a normally distributed uncertainty for a measured reflectivity value. 
While a normally distributed uncertainty is the most common, it may not be accurate in all circumstances, for example, when low numbers of counts are present, a Poisson uncertainty may offer a more accurate description.

@@

If it is not possible to describe the $m$-dimensional distribution using some statistical test and a common distribution type, then confidence intervals can be given. 
Where these are used the percentage of the confidence interval must be defined alongside it. 
Alongside these confidence intervals, it is typically most accurate to give the maximum probability value for the parameter, rather than the numerical mean which may sit in a region of low probability.
When reporting these quantiles of interest, we should assess how much precision we prescribe to them, which is typically achieved by defining some Monte Carlo standard error (MCSE) \cite{vehtari_rank_2021}. 
This is the variance that would be observed should the sampling process be repeated. 
Theres a range of approaches to compute the MCSE, including the \texttt{mcse} method from the \texttt{ArviZ} package \cite{kumar_arviz_2019}.
It is important to check that the MCSE is small enough to report the level of precision desired for a given parameter.

Regardless of how the chain is shared, as a component of a fully reproducible analysis, the author should give details of the software packages, scripts, and data used to produce the analysis and any random number seeds that were defined. 
This means that if the chain is not available, the reader can rerun the sampling and replicate the results. 
Included in this is information regarding specific version numbers for different software packages, as these can create irreplicable results between version numbers.

>>>>

% Define document class
\documentclass[reprint,superscriptaddress,aps,amsmath]{revtex4-2}

@@

Reflectometry analysis can be described, in the most simplistic terms, as a comparison and refinement of a model based on some parameters, $\mathbf{x}$, to reproduce some reflectivity data set, $\mathbf{D}$. 
This refinement process involves comparing the model to the data and calculating some goodness-of-fit value or likelihood, $p(\mathbf{D} | \mathbf{x})$, and modifying the model to optimize the goodness-of-fit or maximise the likelihood.
A commonly used goodness-of-fit parameter is the $\chi^2$ parameter which is found as \cite{nelson_refnx_2019}, 
%
\begin{equation}
    \chi^2 = \sum_{q=q_{\text{min}}}^{q_{\text{max}}}{\bigg[\frac{(R(q) - R(q)_{\text{m}})}{\sigma_R(q)}\bigg]^2}, 
    \label{equ:chi2}
\end{equation}
%
where, $R(q)$ and $R(q)_{\text{m}}$ are the measured and modelled reflectivity at a given $q$, the measured momentum transfer vector, and $\sigma_R(q)$ is the uncertainty associated with the measured reflectivity at each $q$ point. 
Under an assumption of normally distributed residuals $R(q) - R(q)_{\text{m}} \sim \mathcal{N}(0, \sigma_R(q))$, the likelihood is related to the $\chi^2$ variable in the following way:
%
\begin{equation}
    \ln[p(\mathbf{D} | \mathbf{x})] = -\frac{1}{2} \bigg(\chi^2 + \sum_{q=q_{\text{min}}}^{q_{\text{max}}}\ln{\big[2\pi\sigma_R(q)^2\big]}\bigg).
    \label{equ:likelihood}
\end{equation}
%
The input for this refinement process is the model and some initial parameter values, which may be an absolute value or some parameter range, depending on the refinement algorithm.
While for Bayesian sampling processes, the input is a probability distribution for each parameter, known as the prior.
The output is a set of values for $\mathbf{x}$, potentially with associated error bars, where these are present they typically describe a standard deviation from the mean of a Gaussian probability distribution. 
A Bayesian sampling process gives a probability distribution, the posterior, the defines the relative likelihood of different values of each parameter, from this we can report statistical measures, e.g. mode/mean/median.
This process implicitly assumes that the data is completely reduced, accounting for all experimental parameters, uncertainties are accurately described and the model can accurately describe the data.

@@

The optimised parameters from the minimisation algorithm, which depend on the particular algorithm, often include some statistical uncertainty.
This uncertainty comes from an assumption of normally distributed parameters \cite{bevington_data_2002}, however, Bayesian sampling approaches make no assumption of an underlying statistical distribution.
How these statistical uncertainties are found is beyond the scope of this work, but it is important to acknowledge that this uncertainty typically assumes that the probability distribution of the parameter is Gaussian in nature. 
This probability distribution is either the partial likelihood or posterior, the latter when some prior is included and Bayes theorem is applied. The posterior describes our understanding of the parameter values given the data that was measured.
When Bayesian modelling is used and the prior is included, the posterior probability is found as, 
%
\begin{equation}
  p(\mathbf{x} | \mathbf{D}) \propto p(\mathbf{D} | \mathbf{x}) p(\mathbf{x}).
  \label{equ:bayes}
\end{equation}
%
Therefore, when Bayesian modelling is performed, the priors and likelihood are of fundamental importance to the results that are obtained (the posterior), and any scientific conclusions that are drawn. 
We note that Equation~\ref{equ:bayes} omits the normalisation term, the Bayesian evidence [$p(\mathbf{D})$], which is discussed in detail elsewhere \cite{sivia_bayesian_1998,mccluskey_general_2020} and can be omitted when model comparison is not begin performed.

@@

Bayes theorem (given in Equation~\ref{equ:bayes}) consists of the product of the prior and the likelihood. 
The former describes our current understanding of the parameters before we conduct some experiments, while the latter describes how well the data is described by the model parameters.
Although Equation~\ref{equ:likelihood} is a common approach to quantify how well the data is described, it assumes a normally distributed uncertainty for the measured reflectivity value.
While a normally distributed uncertainty is the most common, it may not be accurate in all circumstances, for example, when low numbers of counts are present, a Poisson uncertainty may offer a more accurate description, in which case the likelihood function could be changed to that which reflect multi-dimensional Poisson distribution \cite{lass_multinomial_2021}.
Additionally, the likelihood function may be modified by weighting the data at high-$q$, such as by replacing $R(q)$ in Equation~\ref{equ:chi2} with $\log{[R(q)]}$ or $R(q)q^4$.
These aspects make is very important to state explicity how the likelihood for a given model and data is calculated (giving the analysis package used, version number, and if not the default the likelihood option).

@@

If it is not possible to describe the $m$-dimensional distribution using some statistical test and a common distribution type, then confidence intervals can be given. 
Where these are used the percentage of the confidence interval must be defined alongside it. 
Alongside these confidence intervals, it is typically most accurate to give the maximum probability value for the parameter, rather than the numerical mean which may sit in a region of low probability.
When reporting these quantiles of interest, we should assess how much precision we prescribe to them, which is typically achieved by defining some Monte Carlo standard error (MCSE) \cite{vehtari_rank_2021}. 
This is the variability that would be observed should the sampling process be repeated. 
Theres a range of approaches to compute the MCSE, including the \texttt{mcse} method from the \texttt{ArviZ} package \cite{kumar_arviz_2019}.
It is important to check that the MCSE is small enough to report the level of precision desired for a given parameter.

Regardless of how the chain is shared, as a component of a fully reproducible analysis, the author should give details of the software packages, scripts, and data used to produce the analysis and any random number seeds that were defined. 
This means that if the chain is not available, the reader can rerun the sampling and replicate the results. 
Included in this is information regarding specific version numbers for different software packages, as these can create irreplicable results between version numbers. 
We want to emphasize this value of openly sharing the posteriors of some Bayesian sampling approach. 
All posteriors may be utilised as prior probabilities in subsequent analyses, therefore by sharing posteriors we enable improved analysis in future.
